---
title: "Coaching Networks and Career Pathways Analysis"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(igraph)
library(ggraph)
library(tidygraph)
library(lubridate)
```

# Overview

This analysis explores coaching networks and career pathways in NCAA women's basketball, focusing on which schools produce future head coaches and identifying coaching "trees."

Your Task: Work through each section, making decisions and identifying potential story angles from the data.

---

# Data Loading

Load the CSV file and check out the data to see what you're working with.

```{r load-data}
coaching <- read_csv("https://raw.githubusercontent.com/dwillis/wbb/refs/heads/main/ncaa/coaches/coaching_histories_merged.csv")

cat("\nDataset summary:\n")
cat("Total records:", nrow(coaching), "\n")
cat("Unique coaches:", n_distinct(coaching$coach), "\n")
cat("Unique schools:", n_distinct(coaching$college), "\n")
cat("Division I records:", sum(coaching$division == "I", na.rm = TRUE), "\n")
```

---

# Analysis 1: Head Coach Production Pipeline

## DECISION POINT 1: Defining "Future Head Coaches"

**Question for you**: Should we focus only on coaches who became D1 head coaches, or also include those who became head coaches at D2, D3, or other levels? How does this choice affect the story? Does it matter if someone became a D3 head coach vs. stayed in D1 as an assistant?

**Your choice** (modify the code below):

```{r head-coach-pipeline}
# STUDENT DECISION: Modify this filter as needed
# Option A: Only D1 head coaches (current setting)
# Option B: All NCAA head coaches (remove division filter)
# Option C: D1 + D2 only (division %in% c("I", "II"))

d1_head_coaches <- coaching |>
  filter(position_title_standardized == "Head Coach") |>  # MODIFY THIS LINE
  distinct(coach) |>
  pull(coach)

cat("Number of head coaches identified:", length(d1_head_coaches), "\n")
```


## Finding Previous Positions

```{r previous-positions}
# Find their previous positions (before becoming head coach)
future_hc_positions <- coaching |>
  filter(coach %in% d1_head_coaches, position_title_standardized != 'Head Coach') |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    is_head_coach = position_title_standardized == "Head Coach",
    first_hc_year = min(start_year[is_head_coach], na.rm = TRUE),
    before_hc = start_year < first_hc_year | is.na(first_hc_year)
  ) |>
  ungroup() 

cat("Total positions held by future head coaches:", nrow(future_hc_positions), "\n")
```

## DECISION POINT 2: Which Division Matters for Training?

**Question for you**: When looking at which schools produce future head coaches, should we only count their D1 experience, or also their D2/D3/other experience?

```{r production-schools}
# STUDENT DECISION: Modify the division filter
# Current: Only counting D1 positions
# Alternative: Remove division filter to count all positions

hc_production_schools <- future_hc_positions |>
  filter(before_hc) |>  # MODIFY THIS LINE IF NEEDED
  group_by(college_clean, position_title_standardized) |>
  summarize(
    future_hcs = n_distinct(coach),
    .groups = "drop"
  ) |>
  arrange(desc(future_hcs))

cat("\nTop 20 D1 schools producing future head coaches (all positions):\n")
hc_production_schools |>
  group_by(college_clean) |>
  summarize(total_future_hcs = sum(future_hcs), .groups = "drop") |>
  arrange(desc(total_future_hcs)) |>
  head(20) |>
  print(n = 20)
```

**Think about**: Does a stint at a D3 school early in someone's career "count" as part of their development?

## Position Type Breakdown

```{r position-breakdown}
cat("\n\nTop schools by position type:\n")
hc_production_schools |>
  filter(position_title_standardized %in% c("Assistant Coach", "Associate Head Coach")) |>
  group_by(position_title_standardized) |>
  slice_max(future_hcs, n = 10) |>
  arrange(position_title_standardized, desc(future_hcs))
```

### 💡 STORY IDEAS - Section 1

**Based on what you see above, brainstorm 2-3 story angles:**

1. I think digging into a school like Cincinnati, Boston College or East Carolina who hasn't had much NCAA Tournament success but has produced a ton of head coaches would be interesting. 
2. Notre Dame and Maryland have historically been really good women's basketball programs, but they each only have three people who have moved on to be head coaches. I think diving into the loyalty aspect of those two schools would be interesting to see why those assistant coaches have likely passed up head coaching opportunities — or why the assistant coaches aren't getting many looks at open jobs (depending what the response is)
3. 

**Questions to consider:**
- Which school dominates the list? Is this surprising?
- Are there mid-major schools punching above their weight?
- Are there regional patterns?

---

# Analysis 2: Career Ladder Analysis

## DECISION POINT 3: Minimum Position Threshold

**Question for you**: Should we require a minimum number of positions before someone counts as having a "career ladder"? (e.g., exclude people who went straight from 1 assistant job to head coach?)

```{r career-ladder}
# STUDENT DECISION: Set minimum number of positions
MIN_POSITIONS <- 2  # CHANGE THIS NUMBER (try 2, 3, or keep at 1)

# Sequence of positions before becoming head coach
career_sequences <- future_hc_positions |>
  filter(before_hc, division == "I") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  summarize(
    positions_held = list(position_title_standardized),
    n_positions = n(),
    final_position = last(position_title_standardized),
    .groups = "drop"
  ) |>
  filter(n_positions >= MIN_POSITIONS)  # APPLIES YOUR THRESHOLD

cat("Coaches meeting threshold:", nrow(career_sequences), "\n")

cat("\nMost common final position before becoming D1 head coach:\n")
career_sequences |>
  count(final_position, sort = TRUE) |>
  mutate(pct = n / sum(n) * 100) |>
  head(10)

# Average number of positions held before head coach role
cat("\n\nAverage D1 positions held before becoming head coach:",
    round(mean(career_sequences$n_positions, na.rm = TRUE), 1))
```

## Common Career Pathways

```{r pathways}
# STUDENT DECISION: How many positions should we examine?
PATHWAY_LENGTH <- 3 # CHANGE THIS (2, 3, 4, or 5)

cat("\n\nCommon", PATHWAY_LENGTH, "-position pathways to head coach:\n")
career_sequences |>
  filter(n_positions == PATHWAY_LENGTH) |>
  mutate(pathway = map_chr(positions_held, ~paste(.x, collapse = " → "))) |>
  count(pathway, sort = TRUE) |>
  head(15) |>
  print(n = 15)
```

### 💡 STORY IDEAS - Section 2

**Based on the career ladder data, brainstorm 2-3 story angles:**

1. A piece on the person who went from two years as an assistant coach to a recruiting coordinator would be interesting. I'm curious to see why she technically went down in the coaching staff tree instead of potentially pursuing a head coaching role elsewhere or remaining in her current role
2. I think a piece on the average tenure of an assistant coach would be cool. There were 23 people who remained an assistant coach for four years, which I would consider a lengthy time at that role. You can even find extreme cases of assistant coaches who were quickly promoted to head coach at their current program or at a different school.
3. 

**Questions to consider:**
- Are there "fast track" coaches who skipped steps?
- Does the Associate Head Coach position matter, or do people jump from Assistant to Head?
- How important is the "Staff" level as a starting point?

---

# Analysis 3: Coaching Trees and Mentorship Networks

## DECISION POINT 4: Time Overlap Requirements

**Question for you**: How should we define "worked under"? Should we require they overlapped for at least one full year, or is any overlap sufficient?

```{r coaching-trees}
# STUDENT DECISION: Set minimum overlap threshold
MIN_OVERLAP_YEARS <- 2  # CHANGE THIS (0 = any overlap, 1, 2, 3 for stricter definition)

# Create edges: assistant coaches who worked under head coaches
coaching_relationships <- coaching |>
  filter(division == "I") |>
  select(coach, college_clean, position_title_standardized, start_year, end_year) |>
  # For each school/timeframe, find head coaches
  left_join(
    coaching |>
      filter(division == "I", position_title_standardized == "Head Coach") |>
      select(hc_name = coach, college_clean, 
             hc_start = start_year, hc_end = end_year),
    by = "college_clean",
    relationship = "many-to-many"
  ) |>
  # Keep only overlapping tenures (assistant worked under this head coach)
  filter(
    !is.na(hc_start),
    coach != hc_name,
    position_title_standardized %in% c("Assistant Coach", "Associate Head Coach"),
    # Check for time overlap
    start_year <= coalesce(hc_end, 2025),
    coalesce(end_year, 2025) >= hc_start
  ) |>
  # STUDENT DECISION: Add minimum overlap requirement
  mutate(
    overlap_years = pmin(coalesce(end_year, 2025), coalesce(hc_end, 2025)) - 
                    pmax(start_year, hc_start)
  ) |>
  filter(overlap_years >= MIN_OVERLAP_YEARS) |>  # APPLIES YOUR THRESHOLD
  distinct(coach, hc_name, college_clean)

cat("Coaching relationships identified:", nrow(coaching_relationships), "\n")
```

## Top Mentors

```{r mentors}
# Most prolific mentors (head coaches who produced assistants who became HCs)
prolific_mentors <- coaching_relationships |>
  filter(coach %in% d1_head_coaches) |>
  count(hc_name, sort = TRUE) |>
  rename(mentor = hc_name, proteges_who_became_hc = n)

cat("\nTop 20 mentors (head coaches whose assistants became D1 head coaches):\n")
print(prolific_mentors |> head(20), n = 20)
```

### 💡 STORY IDEAS - Section 3

**Based on the coaching trees data, brainstorm 2-3 story angles:**

1. UConn has been a national powerhouse under Geno Auriemma, but yet he really hasn't mentored many people to become head coaches. I would be curious to see how long people on his current staff have worked at the school and see if his assistant coaches are just content winning national championships at UConn or if they are hoping to move themselves up to the top of the career ladder.
2. Brenda Frese has mentored seven future head coaches, including recently Karen Blair. I think a story on the types of assistant coaches Brenda looks for would be cool. Whatever her method is has been working.
3. 

**Questions to consider:**
- Who are the most prolific "mentor" coaches?
- Are there surprising names?
- Do certain coaches have "coaching trees" at multiple schools?
- Could you profile a specific coaching tree? Which one?

## DECISION POINT 5: Network Visualization Scope

**Question for you**: How many top mentors should we include in the visualization? Too many = cluttered, too few = missing important connections.

```{r network-visualization, fig.width=12, fig.height=10}
# STUDENT DECISION: Number of top mentors to visualize
N_MENTORS_TO_SHOW <- 15  # CHANGE THIS (try 10, 15, 20, or 25)

# Create network of top mentors (for visualization clarity)
top_mentors <- prolific_mentors |>
  head(N_MENTORS_TO_SHOW) |>
  pull(mentor)

network_edges <- coaching_relationships |>
  filter(
    hc_name %in% top_mentors,
    coach %in% d1_head_coaches
  ) |>
  select(from = hc_name, to = coach, school = college_clean)

# Build network graph
coaching_network <- as_tbl_graph(network_edges, directed = TRUE)

# Plot
ggraph(coaching_network, layout = 'fr') +
  geom_edge_link(aes(alpha = 0.3), arrow = arrow(length = unit(2, 'mm'))) +
  geom_node_point(size = 5, color = "steelblue") +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  theme_void() +
  labs(
    title = "Coaching Trees: Top Mentors and Their Protégés",
    subtitle = paste("Top", N_MENTORS_TO_SHOW, "D1 head coaches who mentored future D1 head coaches")
  )
```

---

# Analysis 4: School-to-School Coaching Pipelines

```{r school-pipelines}
# Track coach movement between schools
coach_movements <- coaching |>
  filter(division == "I") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    next_school = lead(college_clean),
    next_position = lead(position_title_standardized),
    next_start = lead(start_year)
  ) |>
  filter(!is.na(next_school), college_clean != next_school) |>
  ungroup()

# Most common school-to-school transitions
cat("\nTop 30 school-to-school coaching pipelines:\n")
coach_movements |>
  count(college_clean, next_school, sort = TRUE) |>
  rename(from_school = college_clean, to_school = next_school, coaches = n) |>
  head(30) |>
  print(n = 30)
```

## Schools that Export to Head Coaching

```{r exports}
# Schools that "export" coaches to D1 head coaching jobs
coach_exports <- coach_movements |>
  filter(next_position == "Head Coach") |>
  count(college_clean, sort = TRUE) |>
  rename(school = college_clean, coaches_to_hc = n)

cat("\n\nTop 20 schools that export coaches to D1 head coaching positions:\n")
print(coach_exports |> head(20), n = 20)
```

### 💡 STORY IDEAS - Section 4

**Based on school-to-school movement patterns, brainstorm 2-3 story angles:**

1. Four coaches going from Maryland to Rice is super interesting. I would imagine three people followed Tina Langley from Maryland, but I'm curious to see what kind of promotions they received when they moved over.
2. Brenda Frese helped seven people become head coaches, but yet Maryland isn't anywhere to be seen on the coaches to head coaches ranking. Why is that the case? Are Maryland coaches just not getting the immediate look to be head coaches elsewhere?
3. 

**Questions to consider:**
- Are there strong regional pipelines (e.g., schools from the same conference feeding each other)?
- Do assistant coaches at top programs tend to move to head coaching at mid-majors first?
- Are there unusual cross-country connections?
- Which schools are "feeders" vs. "destinations"?

---

# Analysis 5: Conference-Level Analysis

## DECISION POINT 6: Conference Focus

**Question for you**: Should we analyze all conferences equally, or focus on major/Power 5 conferences? Should we combine some smaller conferences?

```{r conference-analysis}
# STUDENT DECISION: Define which conferences to highlight
# Option A: Analyze all conferences (current)
# Option B: Create a "Power 5" vs "Mid-Major" vs "Small" grouping
# Option C: Focus only on specific conferences

# Conferences producing future head coaches
conference_production <- future_hc_positions |>
  filter(before_hc, division == "I") |>
  count(conference, position_title_standardized) |>
  arrange(desc(n))

cat("\nTop conferences producing future D1 head coaches:\n")
conference_production |>
  group_by(conference) |>
  summarize(total = sum(n), .groups = "drop") |>
  arrange(desc(total)) |>
  head(15) |>
  print(n = 15)
```

## Conference Loyalty

```{r loyalty}
# Conference loyalty: coaches who stay within conference
conference_loyalty <- coach_movements |>
  left_join(
    coaching |> 
      filter(division == "I") |>
      select(coach, college_clean, conference),
    by = c("next_school" = "college_clean", "coach" = "coach"),
    relationship = "many-to-many"
  ) |>
  filter(!is.na(conference.x), !is.na(conference.y)) |>
  mutate(same_conference = conference.x == conference.y) |>
  group_by(conference.x) |>
  summarize(
    total_moves = n(),
    same_conference_moves = sum(same_conference),
    loyalty_rate = same_conference_moves / total_moves * 100,
    .groups = "drop"
  ) |>
  arrange(desc(loyalty_rate))

cat("\n\nConference retention rates (coaches staying within conference):\n")
print(conference_loyalty |> head(15), n = 15)
```

### 💡 STORY IDEAS - Section 5

**Based on conference patterns, brainstorm 2-3 story angles:**

1. While the Big Ten has been considered the second-best conference behind the SEC, there aren't many people that move within the Big Ten. I would like to examine why that is the case.
2. There have been 326 total moves from ACC schools, but only 35 have landed at other ACC schools. What is the most common destination (conference-wise) for those 291 people.
3. 

**Questions to consider:**
- Do certain conferences "develop" more coaches than others?
- Is there high loyalty within conferences, or do coaches move around?
- Are there conferences that import talent vs. export it?

---

# Analysis 6: Time to Head Coach Role

## DECISION POINT 7: Career Start Definition

**Question for you**: When calculating "time to head coach," should we count ALL positions (including high school, community college) or only college-level positions?

```{r time-to-hc}
# STUDENT DECISION: Filter on category
# Current: All positions included
# Alternative: Add filter like "category == 'College'" to exclude high school, etc.

time_to_hc <- coaching |>
  filter(coach %in% d1_head_coaches) |>
  filter(category == "College") |>  # UNCOMMENT THIS LINE TO EXCLUDE NON-COLLEGE
  arrange(coach, start_year) |>
  group_by(coach) |>
  summarize(
    first_position_year = min(start_year, na.rm = TRUE),
    first_hc_year = min(start_year[position_title_standardized == "Head Coach" & 
                                     division == "I"], na.rm = TRUE),
    years_to_hc = first_hc_year - first_position_year,
    .groups = "drop"
  ) |>
  filter(is.finite(years_to_hc), years_to_hc >= 0)

cat("\nTime to D1 head coach position:\n")
cat("Mean:", round(mean(time_to_hc$years_to_hc, na.rm = TRUE), 1), "years\n")
cat("Median:", median(time_to_hc$years_to_hc, na.rm = TRUE), "years\n")
cat("Min:", min(time_to_hc$years_to_hc, na.rm = TRUE), "years\n")
cat("Max:", max(time_to_hc$years_to_hc, na.rm = TRUE), "years\n")

# Distribution
ggplot(time_to_hc, aes(x = years_to_hc)) +
  geom_histogram(binwidth = 2, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = median(time_to_hc$years_to_hc, na.rm = TRUE), 
             linetype = "dashed", color = "red", size = 1) +
  labs(
    title = "Years to First D1 Head Coaching Position",
    subtitle = "From first coaching role to D1 head coach",
    x = "Years",
    y = "Count of coaches"
  ) +
  theme_minimal()
```

### 💡 STORY IDEAS - Section 6

**Based on career timeline data, brainstorm 2-3 story angles:**

1. There are over 40 coaches who have landed D1 head coaching jobs after no years of coaching at the D1, D2, or D3 level — likely from NJCAA schools. Why are community college and JUCO colleges finding more immediate success at becoming head coaches in D1 versus assistant coaches already at that level?
2. There is one person who waited over 30 years coaching at the Division 1 level to land a head coaching job. Who is that person and what is his story?
3. 

**Questions to consider:**
- What's the typical timeline to becoming a D1 head coach?
- Who are the outliers (very fast or very slow)?

---

# Analysis 7: Elite Program "Finishing Schools"

## DECISION POINT 8: Defining "Elite Programs"

**Question for you**: Which schools should count as "elite" or "blue blood" programs? 
This is subjective - modify the list based on your knowledge or research.

```{r finishing-schools}
# STUDENT DECISION: Modify this list of top programs
# Consider: recent success, historical success, resources, national profile

top_programs <- c("UConn", "Tennessee", "Stanford", "Notre Dame", "Baylor",
                  "South Carolina", "LSU", "Texas", "Duke", "UCLA",
                  "Louisville", "Maryland", "Ohio St.", "Iowa")
# ADD OR REMOVE SCHOOLS AS YOU SEE FIT

cat("Analyzing", length(top_programs), "programs\n")

# Assistant coaches from top programs who became head coaches
finishing_school_analysis <- future_hc_positions |>
  filter(
    before_hc,
    division == "I",
    college_clean %in% top_programs,
    position_title_standardized %in% c("Assistant Coach", "Associate Head Coach")
  ) |>
  group_by(college_clean) |>
  summarize(
    future_head_coaches = n_distinct(coach),
    .groups = "drop"
  ) |>
  arrange(desc(future_head_coaches))

cat("\n'Finishing schools' - top programs producing future head coaches:\n")
print(finishing_school_analysis, n = Inf)

# Average years spent at top program before becoming HC
years_at_top_programs <- future_hc_positions |>
  filter(
    before_hc,
    division == "I",
    college_clean %in% top_programs,
    position_title_standardized %in% c("Assistant Coach", "Associate Head Coach")
  ) |>
  mutate(years_at_school = coalesce(end_year, 2025) - start_year) |>
  group_by(college_clean) |>
  summarize(
    avg_years = mean(years_at_school, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(avg_years))

cat("\n\nAverage years spent at top programs (for future HCs):\n")
print(years_at_top_programs, n = Inf)
```

### 💡 STORY IDEAS - Section 7

**Based on elite program patterns, brainstorm 2-3 story angles:**

1. Iowa staff personnel spend by far the most time at the school before becoming a head coach — I would imagine most of that is under Lisa Bluder. What about her keeps coaches around the program for a lengthy amount?
2. Why are schools like Maryland and Tennessee having a much greater success immediately producing head coaches versus recent long-time powerhouses UConn and South Carolina?
3. 

**Questions to consider:**
- Do the biggest brand names produce the most head coaches?
- Are there "hidden gems" - programs that punch above their weight?
- Does time spent at elite programs matter (short stint vs. long apprenticeship)?
- Are there programs that hoard talent vs. those that develop and release it?

---

# Analysis 8: "Stepping Stone" Schools

```{r stepping-stones}
# Head coaches who left for another D1 head coaching job
stepping_stone_moves <- coaching |>
  filter(division == "I", position_title_standardized == "Head Coach") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    next_school = lead(college_clean),
    next_position = lead(position_title_standardized),
    moved_to_hc = next_position == "Head Coach" & !is.na(next_school)
  ) |>
  filter(moved_to_hc) |>
  ungroup()

cat("\nSchools most commonly serving as stepping stones (HC → HC moves):\n")
stepping_stone_moves |>
  count(college_clean, sort = TRUE) |>
  rename(school = college_clean, hc_departures = n) |>
  head(20) |>
  print(n = 20)

# Most common destination schools
cat("\n\nMost common destination schools for head coaches:\n")
stepping_stone_moves |>
  count(next_school, sort = TRUE) |>
  rename(destination = next_school, hc_arrivals = n) |>
  head(20) |>
  print(n = 20)
```

### 💡 STORY IDEAS - Section 8

**Based on stepping stone patterns, brainstorm 2-3 story angles:**

1. Why is George Washington such a destination school for head coaches? I'm not very in the loop of its women's basketball program, but I haven't really heard of any NCAA Tournament success its had. 
2. South Dakota churns out head coaches who likely move up to Power Four coaching jobs. What are they looking for in head coaches that has translated well to get poached by bigger schools?
3. 

**Questions to consider:**
- Which schools can't keep their successful head coaches?
- What makes a school a "destination" vs. a "stepping stone"?
- Are there schools that serve as consistent "first head coaching jobs"?
- Can you identify a "ladder" of schools (small to medium to major)?

---

# Analysis 9: Position Mobility Analysis

```{r position-mobility}
# Years spent in each position before moving up
position_duration <- coaching |>
  filter(division == "I") |>
  mutate(years_in_position = coalesce(end_year, 2025) - start_year) |>
  group_by(position_title_standardized) |>
  summarize(
    n_positions = n(),
    avg_duration = mean(years_in_position, na.rm = TRUE),
    median_duration = median(years_in_position, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(n_positions))

cat("\nAverage years in each position type:\n")
print(position_duration, n = Inf)

# Promotion rates by position
promotions <- coaching |>
  filter(division == "I") |>
  arrange(coach, start_year) |>
  group_by(coach) |>
  mutate(
    next_position = lead(position_title_standardized),
    promoted = case_when(
      position_title_standardized == "Staff" & 
        next_position %in% c("Assistant Coach", "Associate Head Coach", "Head Coach") ~ TRUE,
      position_title_standardized == "Assistant Coach" & 
        next_position %in% c("Associate Head Coach", "Head Coach") ~ TRUE,
      position_title_standardized == "Associate Head Coach" & 
        next_position == "Head Coach" ~ TRUE,
      TRUE ~ FALSE
    )
  ) |>
  filter(!is.na(next_position)) |>
  group_by(position_title_standardized) |>
  summarize(
    total_moves = n(),
    promotions = sum(promoted),
    promotion_rate = promotions / total_moves * 100,
    .groups = "drop"
  ) |>
  arrange(desc(promotion_rate))

cat("\n\nPromotion rates by position:\n")
print(promotions, n = Inf)
```

### 💡 STORY IDEAS - Section 9

**Based on mobility and promotion patterns, brainstorm 2-3 story angles:**

1. Why are recruiting coordinators never getting promoted? I feel like that experience would eventually make for a pretty successful head coach, just being able to make connections with recruits and convince them to join the program
2. Interim head coaches are usually ones that have been in the program for awhile, and are experienced on the coaching staff. Why are they never able to actually get the head coaching job?
3. 

**Questions to consider:**
- Which positions are "dead ends" vs. "springboards"?
- Do people spend more time in certain positions?
- Is there a "right" amount of time to spend before moving up?

---

# FINAL REFLECTION: Your Best Story Ideas

Review all your story ideas from each section. Which 3 would you pursue as full stories?

**Story #1:** Brenda Frese has mentored seven future head coaches, including recently Karen Blair. I think a story on the types of assistant coaches Brenda looks for would be cool. Whatever her method is has been working.

**Why this story:** Brenda Frese usually doesn't have many men on her staff and has entirely shifted to an all woman's staff in recent years. But since there aren't many women head coaches across the country, I think it would interesting to see why she is having a greater success putting women's assistant coaches on the map.

---

**Story #2:** There is one person who waited over 30 years coaching at the Division 1 level to land a head coaching job. Who is that person and what is their story?

**Why this story:** Coaching for 30 years at the Division I level is very rare, especially as just an assistant coach. I'd like to get an understanding of why he waited 30 years to take a head coaching job — if he had offers throughout his tenure — or if it was something they had always wanted to do and took a head coaching job a couple years before retirement just to accomplish it.

---

**Story #3:** Interim head coaches are usually ones that have been in the program for awhile, and are experienced on the coaching staff. Why are they never able to actually get the head coaching job?

**Why this story:** When interim head coaches get their chance to lead the team after the head coach is fired, I feel like there has to be one that finds success and revitalizes the program. I'd like to know why that is a thing in women's basketball, where of 41 interim head coaches, none of them ever got the permanent head coach tag.

---

# Additional Analysis Ideas to Explore

Now that you've worked through these analyses, what other questions could this data answer?

**Your ideas:**

1. How long should you roughly have to wait on a coaching staff to eventually become a head coach?
2. What the most successful schools are at retaining assistants and others on the coaching staff for such a long time?
3. If you are looking to become a head coach, what conference should you work in to have the best chance of leading a program?
