## NCAA Football Attendance 

Let's load the tidyverse.

```{r}
library(tidyverse)
```

And read in the attendance data:

```{r}
attendance <- read_csv("https://dwillis.github.io/sports-data-files/attendance.csv")
```

Using the tidyverse's `pivot_longer` function, turn the currently wide data into longer data and save it as a new dataframe called `attendance_longer`:

```{r}
attendance_longer <- attendance |> pivot_longer(cols = starts_with("20"), names_to = "Year", values_to = "Attendance")
```

Let's try to figure out whether football attendance has returned to "normal" by 2024, defining "normal" as 2019 figures. For that we'll need to isolate those two years using `filter()` and get rid of any NAs:

```{r}
attendance_two_years <- attendance_longer |> filter(Year == 2019 | Year == 2024) |>  na.omit()
```

Now, in order to calculate a percentage change between those two years, we'll need to go back to wider data and again ditch teams where we don't have both years:

```{r}
attendance_wide <- attendance_two_years |> pivot_wider(names_from = Year, values_from = Attendance) |> na.omit()
```

Finally, we'll add a percentage change column to that dataframe (remember, it's (New-Old)/Old) and arrange it to see the results.

```{r}
attendance_wide <- attendance_wide |> mutate(percent_change = (`2024`-`2019`)/`2019` * 100) |> 
arrange (percent_change)

```

Some of these clearly are influenced by performance on the field. Below, describe the most interesting results (positive or negative) and what you think might be going on with them. Feel free to do some googling and describe what other data would be helpful to clarify what the best stories here might be. What surprises you about these results?

I think the biggest story is what is transpiring with the Akron football team. To go from an attendance of 107,752 in 2019 to only 31,872 is a drastic change. While Akron played one less game at home last year than they did in 2019, that doesn't explain that big drop-off. In 2019, Akron also didn't win a football game so I would expect that number to be lower than it actually was. Maybe the continuous coaching change has made an impact — they've had two new coaches in that time frame — but that still doesn't make much sense with the attendance fall off. 



