---
title: "wbb"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
wbb_stats <- read_csv("https://dwillis.github.io/sports-data-files/wbbstats1525.csv")
```

```{r}
wbb_new_stats <- wbb_stats |> 
  mutate (
    ppg = ForPoints/Games,
    Opp_ppg = OppPoints/Games,
    TotalRebounds_per_game = TotalRebounds/Games,
    Steals_per_game = Steals/Games,
    Turnovers_per_game = Turnovers/Games,
    )
```

```{r}
teamzscore <- wbb_new_stats |> 
  mutate(
    Pointszscore = as.numeric(scale(ppg, center = TRUE, scale = TRUE)),
    OppPointszscore = as.numeric(scale(Opp_ppg, center = TRUE, scale = TRUE)) *-1,
    FGzscore = as.numeric(scale(FieldGoalPCT, center = TRUE, scale = TRUE)),
    RebZscore = as.numeric(scale(TotalRebounds_per_game, center = TRUE, scale = TRUE)),
    TurnoversZscore = as.numeric(scale(Turnovers_per_game, center = TRUE, scale = TRUE)) *-1,
    StealsZscore = as.numeric(scale(Steals_per_game, center = TRUE, scale = TRUE)),
    TotalZscore = Pointszscore + OppPointszscore + FGzscore + RebZscore + TurnoversZscore + StealsZscore
  )
```

The highest performing teams are either from UCONN, South Carolina or Baylor, which makes the most sense given their periods of dominance in women's college basketball. The three programs have combined for seven of the last ten national championships and continuously produced quality teams. The biggest surprise to me is how far down the 2024 Iowa team is, when they made the run to the national championship game. With Caitlin Clark leading the charge, I figured they would get a big boost from the offensive metrics. The 2017 Maryland team finds their way towards the top of the list and the 2020 Maryland team isn't too far behind.

```{r}
teamzscore |> 
  filter(School == "Maryland") |> 
  arrange(desc(TotalZscore)) |>
  select(Season, School, TotalZscore)
```

The top two Maryland teams are the 2017 and 2020 ones. The 2017 team made a run to the Sweet 16 and likely would have advanced further if they didn't run into a red-hot Oregon 10-seed. What is most surprising to me is that the 2020 team is second on the list. If not for the pandemic, that 2020 team was destined for at least the Final Four, maybe even a national championship. The depth and talent on that roster is without a doubt the best Brenda Frese has had.

A story idea just focused on Maryland would be how last year's team over performed relative to the other Maryland teams ahead of it on the list. While they may not have been the most talented team, making the Sweet 16 and nearly beating the eventual runners-up is impressive. Last year's team just found ways to win games.

An overall story idea is comparing how the national championship winners did in this metric versus the team they played in the championship game. I think it would be interesting to see if the "best" team really won or if the winner just performed really well in that game or the loser just had a really bad game.

### Z-Score Writeup

Basketball head coaches across the country continuously preach the word rebounding to their teams, ingraining it into their players' heads. While good teams can score points in a hurry and defend for 40 minutes, crashing the glass is often a controllable that separates good from elite teams.

That is the identity of Troy women's basketball under head coach Chanda Rigby. While the Trojans may not have overpowering size like UCLA's Lauren Betts to gather rebounds over opposing defenders, they are one of the most well-positioned teams when shots go up. It shows.

Of the 3,880 Division I teams from 2015-2025, Troy makes up four of the top five rebounding teams and two more find their way into the top-10. Outside of the 2023 South Carolina team — led by Aliyah Boston, one of the best rebounders in women's college basketball history — the Trojans have controlled the glass in recent years. Those six teams combine to average just over 48 rebounds a game, including a pair of seasons north of 50.

And this philosophy has translated to success. Despite failing to produce an NCAA Tournament win in her tenure thus far — something that is hard for mid-majors to do — Troy has remained a powerhouse in the Sun Belt conference, contending for the league's auto bid every season. The Trojans have totaled right around 22 wins a season, with a 17-win year in 2022-23 dragging the average significantly down.

Because of these elite rebounding seasons, Troy is typically able to jack up more shots than its opponents. The combination of second-chance opportunities and limiting other teams to just one shot per possession is a formula for great results. 