---
title: "Women's Soccer Data"
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1G1AjH2iS6kj3kKHCRi0XVy5Lcm-rahJkeTAToVh9J1g/edit?gid=0#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2025 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:


```{r}
matches_2025 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2025.csv") |> mutate(diff = team_score - opponent_score, total_goals = goals + defensive_goals)
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv") |> mutate(diff = team_score - opponent_score, total_goals = goals + defensive_goals)
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2025)
glimpse(matches_2025)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1. How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal.

```{r}
matches_2025 |>
# filter(team == "Maryland Terrapins, Terps") |>
  group_by(team) |> 
  summarise(
    shots = sum(sh_att),
    shots_on_goal = sum(so_g),
    shots_on_goal_pct = (shots_on_goal/shots) * 100
  ) |> 
  arrange(desc(shots_on_goal_pct))
```

2. What about the same shot stats from last season's first four games? You'll need to load the matches from 2024 - just change the season in the URL - and filter by date to find the first four games.

```{r}
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv")
matches_2024 |>
  filter(team == "Maryland Terrapins, Terps", date < "2024-08-26") |>
  summarise(
    shots = sum(sh_att),
    shots_on_goal = sum(so_g),
    shots_on_goal_pct = (shots_on_goal/shots) * 100
  )
```


3. Maryland started 2022 with 4 draws. Have any other teams had as many this season?

```{r}
matches_2025 |>
  filter(outcome == "Draw") |>
  group_by(team) |>
  summarise(draws = n()) |>
   filter (draws >=4) |>
   arrange(desc(draws))
```


4. How many total corner kicks has Maryland had so far this season and how many have its opponents have? Calculate the difference between the two totals for every team using group_by, summarize and mutate, then filter to find Maryland's totals.

```{r}
matches_2025 |>
  group_by(team) |>
  summarise(
    corners_kicks = sum(corners),
    opp_corner_kicks = sum(defensive_corners)
  ) |>
  mutate(corner_diff = corners_kicks - opp_corner_kicks) |>
  filter(team == "Maryland Terrapins, Terps")
```

5. What's the most interesting thing you've seen from these questions?

I find it really interesting that there are two teams (Cal Poly Mustangs and Drexel Dragons) with at least four ties already this season. Ties in soccer are relatively common but to have over half of your games be ties at this point in the season is crazy. 

```{r}
swac_teams <- c("Alabama A&M Bulldogs", "Alabama St. Lady Hornets", "Alcorn Lady Braves", "Ark.-Pine Bluff Lady Lions", "Grambling Lady Tigers", "Jackson St. Lady Tigers", "Mississippi Val. Devilettes", "Prairie View Lady Panthers", "Southern U. Lady Jaguars", "Texas Southern Lady Tigers")

swac_matches <- matches_2025 |> 
  filter(team %in% swac_teams)

matches_2025 |> 
  summarise(
    mean_diff = mean(diff),
    mean_opp_goals = mean(defensive_goals),
    mean_sog = mean(so_g),
    mean_opp_sog = mean(defensive_so_g)
    )

swac_matches |> 
  summarise(
    mean_diff = mean(diff),
    mean_opp_goals = mean(defensive_goals),
    mean_sog = mean(so_g),
    mean_opp_sog = mean(defensive_so_g)
    )

big10 <- c("Nebraska Huskers", "Iowa Hawkeyes", "Minnesota Golden Gophers", "Illinois Fighting Illini", "Northwestern Wildcats", "Wisconsin Badgers", "Indiana Hoosiers", "Purdue Boilermakers", "Ohio St. Buckeyes", "Michigan Wolverines", "Michigan St. Spartans", "Penn St. Nittany Lions", "Rutgers Scarlet Knights", "Maryland Terrapins, Terps", "Southern California Trojans", "UCLA Bruins", "Washington Huskies", "Oregon Ducks")

big10_matches <- matches_2025 |> 
  filter(team %in% big10)

big10_matches |> 
  group_by(team) |> 
  summarise(
    mean_diff = mean(diff),
    mean_opp_goals = mean(defensive_goals),
    mean_sog = mean(so_g),
    mean_opp_sog = mean(defensive_so_g)
    )
```
### Significance test
```{r}
matches_2025 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2025.csv") |> mutate(diff = team_score - opponent_score, total_goals = goals + defensive_goals)
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv") |> mutate(diff = team_score - opponent_score, total_goals = goals + defensive_goals)
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv") |> mutate(diff = team_score - opponent_score, total_goals = goals + defensive_goals) |> filter(date < '2024-09-18')
```

t-test
```{r}
t.test(matches_2024$total_goals, matches_2025$total_goals)
```
While the average number of goals scored this season is sightly up (2.75 to 2.90) from last year, the disparity in the number of games played this season hinders our ability to draw any conclusions from the comparison. When we just compare it to this point last season, the average number of goals scored this season is actually slightly down. With the season still in the early stages of conference play, I expect the average number of goals scored this season to go down right around where it finished last year.  

