---
title: "second_presentation"
format: html
---

I collected data of the unbeaten teams heading into the NCAA Tournament over the last three decades. Within the dataset, I included the team name, the year of their unbeaten record, how far they went in the ncaa tournament, their seed, how many games were they unbeaten for and whether it was a women's or men's team.

With this data, I want to see the distribution of the ncaa tournament success each unbeaten team has — both men's and women's. I'm curious to see which programs have done it with a men's team and women's team. I'll dive into the seeding to see if it's actually better not to be the top seed when you are unbeaten.

My question is does it make sense to bring in data of the team who suffered their first loss in their conference tournaments. I think it would be interesting to see if it's actually better to lose just before the NCAA Tournament and get that first defeat out of the way. Maryland this year would become an example of this because they took their first loss of the season in the Big Ten Tournament semifinals and we'll see how far they make it into the tournament (their path is very wide open with so many of the top teams getting eliminated)
```{r}
library(tidyverse)
library(janitor)
unbeaten_data <- read_csv("https://raw.githubusercontent.com/dylanschmidt05/jour479x_fall_2025-main/refs/heads/main/data/UnbeatenTeams.csv") |> clean_names()
```

```{r}
unbeaten_data |> 
  group_by(gender) |> 
  summarise (
    total = n()
  ) |> 
  arrange(desc(total))
```

```{r}
unbeaten_data |> 
  group_by(team) |> 
  summarise (
    total = n()
  )  |> 
  arrange(desc(total))
```

```{r}
mens_unbeaten <- unbeaten_data |> 
  filter (gender == "Men's")

womens_unbeaten <- unbeaten_data |> 
  filter (gender == "Women's")
```

```{r}
mens_unbeaten |> 
  group_by(seed) |> 
  summarise (
    seed_grouping = n()
  ) |> 
  arrange(desc(seed_grouping))

mens_one_seeds <- mens_unbeaten |> 
  filter(seed == "No. 1")

mens_two_seeds <- mens_unbeaten |> 
  filter(seed == "No. 2")

mens_six_seeds <- mens_unbeaten |> 
  filter(seed == "No. 6")

mens_eight_seeds <- mens_unbeaten |> 
  filter(seed == "No. 8")

mens_ten_seeds <- mens_unbeaten |> 
  filter(seed == "No. 10")

mens_thirteen_seeds <- mens_unbeaten |> 
  filter(seed == "No. 13")

mens_unseeded <- mens_unbeaten |> 
  filter(seed == "Unseeded")
```

```{r}
womens_unbeaten |> 
  group_by(seed) |> 
  summarise (
    seed_grouping = n()
  )  |> 
  arrange(desc(seed_grouping))

womens_one_seeds <- womens_unbeaten |> 
  filter(seed == "No. 1")

womens_two_seeds <- womens_unbeaten |> 
  filter(seed == "No. 2")

womens_three_seeds <- womens_unbeaten |> 
  filter(seed == "No. 3")

womens_four_seeds <- womens_unbeaten |> 
  filter(seed == "No. 4")

womens_six_seeds <- womens_unbeaten |> 
  filter(seed == "No. 6")

womens_seven_seeds <- womens_unbeaten |> 
  filter(seed == "No. 7")

womens_thirteen_seeds <- womens_unbeaten |> 
  filter(seed == "No. 13")

womens_unseeded <- womens_unbeaten |> 
  filter(seed == "Unseeded")
```





